---
title: "Homework 6"
author: "Annie Sauer"
date: "10/10/2018"
output:
  html_notebook: default
  html_document: default
---

## Problem 2

```{r}
# Generate Data
set.seed(12345)
y <- seq(from=1, to=100, length.out = 1e8) + rnorm(1e8)

ybar <- mean(y)

# Method A - use a for loop 
system.time({
  sum <- 0
  for (i in 1:length(y)){
    sum <- sum + (y[i] - ybar)^2
  } 
  sst <- sum/(length(y)-1)
})

# Method B - use vector operations
system.time({
  sst2 <- sum((y-ybar)^2)/(length(y)-1)
})
```

Method A ran in 7.534 seconds, but Method B ran in 0.828 seconds.  The vector operations were much faster than the for loop.  Both methods produced an sst value of 817.75.

## Problem 3

```{r}
set.seed(1256)
# Goal - estimate theta_0 and theta_1 using iterations
theta <- as.matrix(c(1, 2), nrow=2) # True theta values
X <- cbind(1, rep(1:10, 10)) # X input (data)
h <- X %*% theta + rnorm(100, 0, 0.2) # Y output (data)

# Initialze vectors to store theta_0 and theta_1 iterations
theta_0 = vector()
theta_1 = vector()
# Choose two starting points (0,0) and (2,3)
theta_0[1] <- 0 
theta_1[1] <- 0
theta_0[2] <- 2
theta_1[2] <- 3
tol <- 0.01 # tolerance
alpha <- 0.01 # step size
i = 1 # starting index of theta vectors
count = 0
max_it = 10000 # maximum iterations
while((abs(theta_0[i+1] - theta_0[i]) > tol) & 
      (abs(theta_1[i+1] - theta_1[i]) > tol)){
  theta_0[i+2] <- theta_0[i] - (alpha/length(X))*sum(theta_0[i]+theta_1[i]*X[,2]-h)
  theta_1[i+2] <- theta_1[i] - (alpha/length(X))*sum((theta_0[i]+theta_1[i]*X[,2]-h)*X[,2])
  count = count + 1
  i = i+1
  if (count > max_it){
    cat("Error: Maximum iterations reached.")
    break
  }
}

# Compare with linear model 
model <- lm(h~0+X)
```

```{r, echo = FALSE}
cat("Iterative Method \n")
cat("Theta_0 = ", theta_0[length(theta_0)], "with error ", theta[1]-theta_0[length(theta_0)], "\n")
cat("Theta_1 = ", theta_1[length(theta_1)], "with error ", theta[2]-theta_1[length(theta_1)], "\n")
cat("Tolerance = ", tol, "\n")
cat("Step Size = ", alpha, "\n\n")

cat("Linear Model \n")
cat("Theta_0 = ", model$coefficients[1], "with error ", theta[1]-model$coefficients[1], "\n")
cat("Theta_1 = ", model$coefficients[2], "with error ", theta[2]-model$coefficients[2], "\n")
```

## Problem 4

To solve
\[\hat{\beta} = (X^TX)^{-1}X^Ty\]
we can rewrite this equation as 
\[(X^TX)\hat{\beta} = X^Ty\]
Then we can use the solve function in R to find $\hat{\beta}$ without inverting any matrices.
```{r}
x <- cbind(1, rnorm(1000, 10, 1))
y <- rnorm(1000, 20, 1)
xTx <- t(x) %*% x
xTy <- t(x) %*% y
solve(xTx, xTy)
```

## Problem 5

```{r}
set.seed(12456) 
    
G <- matrix(sample(c(0,0.5,1),size=16000,replace=T),ncol=10)
R <- cor(G) # R: 10 * 10 correlation matrix of G
C <- kronecker(R, diag(1600)) # C is a 16000 * 16000 block diagonal matrix
id <- sample(1:16000,size=932,replace=F)
q <- sample(c(0,0.5,1),size=15068,replace=T) # vector of length 15068
A <- C[id, -id] # matrix of dimension 932 * 15068
B <- C[-id, -id] # matrix of dimension 15068 * 15068
p <- runif(932,0,1)
r <- runif(15068,0,1)
C<-NULL #save some memory space
```

```{r, echo=FALSE}
cat("A: ", object.size(A), " bytes \n")
cat("B: ", object.size(B), "bytes")
```

```{r}
# Calculate y (no tricks)
system.time({
  y = p + A %*% solve(B) %*% (q-r)
})
```

This code ran in 1110.317 seconds.  Are there ways to make this faster?

Idea 1: Multiplying a matrix by a vector should be faster than multiplying a matrix by another matrix.  We might be able to change the order of operations to increase the speed.

```{r}
# Calculate y (change order of operations)
system.time({
  y = p + A %*% (solve(B) %*% (q-r))
})
```

This code ran in 925.593 seconds - slightly faster.

Idea 2: Inverting $B$ seems to be taking a long time.  If we take the transpose of both sides, we can rewrite the above formula to remove $B^{-1}$ as follows
\[B[(q-r)(q-r)^T]^{-1}(q-r)(y^T-p^T) = A\]
We could then use the solve function to find $y^T-p^T$.  However this still involves inverting $(q-r)(q-r)^T$ which is the same size as $B$.  I suspected this method wouldn't be any faster.  I ran the code but cut it off after 30 minutes when it was confirmed that this equation took longer to run than the original.  

Idea 3: We know that $B$ is symmetric.  Is there any way we could use this to our advantage? 




